Mysis analysis

### 2015-10-12 betaRegression/betaRegressionExample.Rmd

```{r}
library(rstan)
model <- "
// beta Regression
data {
  int<lower=1> N;                   // sample size
  int<lower=1> p;                   // p predictors
  vector<lower=0,upper=1>[N] y;     // response 
  matrix[N, p] X;                   // predictor matrix
}
parameters {
  vector[p] theta;                  // reg coefficients
  real<lower=0> phi;                // dispersion parameter
}
transformed parameters{
  vector[p] beta;
  beta <- theta * 10;               // matt trick if desired
}
model {
  // model calculations
  vector[N] Xbeta;                  // linear predictor
  vector<lower=0,upper=1>[N] mu;    // transformed linear predictor
  vector<lower=0>[N] A;             // parameter for beta distn
  vector<lower=0>[N] B;             // parameter for beta distn
  Xbeta <- X * beta;
  for (i in 1:N) { 
    mu[i] <- inv_logit(Xbeta[i]);   
  }
  A <- mu * phi;
  B <- (1.0 - mu) * phi;
  // priors
  theta ~ normal(0, 1);   
  phi ~ cauchy(0, 5);               // different options for phi  
  //phi ~ inv_gamma(.001, .001);
  //phi ~ uniform(0, 500);          // put upper on phi if using this
  // likelihood
  y ~ beta(A, B);
}
"
write(model, file = "../stanModels/betaRegression-glm.stan")
sm <- stan_model("../stanModels/betaRegression-glm.stan")
```

>>>>

```{r}
library(rstan)
model <- "
// beta Regression
data {
  int<lower=1> N;                   // sample size
  int<lower=1> p;                   // p predictors
  vector<lower=0,upper=1>[N] y;     // response 
  matrix[N, p] X;                   // predictor matrix
}
parameters {
  vector[p] theta;                  // reg coefficients
  real<lower=0> phi;                // dispersion parameter
}
transformed parameters{
  vector[p] beta;
  vector[N] A;             // parameter for beta distn
  vector[N] B;             // parameter for beta distn
  vector[N] mu;    // transformed linear predictor
  vector[N] Xbeta;                  // linear predictor
  beta <- theta * 10;               // matt trick if desired
//  vector<lower=0,upper=1>[N] mu;    // transformed linear predictor
  Xbeta <- X * beta;
  for (i in 1:N) { 
    mu[i] <- inv_logit(Xbeta[i]);   
  }
  A <- mu * phi;
  B <- (1.0 - mu) * phi;
}
model {
  // model calculations
  // priors
  theta ~ normal(0, 1);   
  phi ~ cauchy(0, 5);               // different options for phi  
  //phi ~ inv_gamma(.001, .001);
  //phi ~ uniform(0, 500);          // put upper on phi if using this
  // likelihood
  y ~ beta(A, B);
} generated quantities {
vector[N] log_lik;
  for (n in 1:N){
    log_lik[n] <- beta_log(y[n], A[n], B[n]);
  }
}
"
write(model, file = "~/mysis/stanModels/betaRegression-glm.stan")
sm <- stan_model("~/mysis/stanModels/betaRegression-glm.stan")
```

### 2015-10-12 countModel/countData.Rmd

```{r}
e <- rstan::extract(m, pars = c("beta", "phi"))
par(mfrow = c(2, 3))
for(i in 1:2) {
  if(i == 1) {
#     for(j in 1:6) {
    for(j in 1:5) {
      plot(density(e[[i]][,j]), main = paste("beta", j, sep=""))
    }
  } else if(i == 2) {
    plot(density(e[[i]]), main = "phi")
  }
}
```

>>>>

```{r}
e <- rstan::extract(m, pars = c("beta", "phi"))
library(MASS)
nbmod <- glm.nb(y ~ . -1, data = data.frame(y=count, X))
summary(nbmod)
coefs <- c(nbmod$coefficients, nbmod$theta)
par(mfrow = c(2, 3))
for(i in 1:2) {
  if(i == 1) {
#     for(j in 1:6) {
    for(j in 1:5) {
      plot(density(e[[i]][,j]), main = paste("beta", j, sep=""))
      abline(v = coefs[j], lwd = 3, col = "red")
    }
  } else if(i == 2) {
    plot(density(e[[i]]), main = "phi")
    abline(v = coefs[dim(e[[i]])[2] + 1], lwd = 3, col = "red")
  }
}
```

### 2015-10-12 juvenileRatioModel/juvenileRatioModel.Rmd

---
title: "negativeBinomial"
author: "John Tipton"
date: "August 14, 2015"
output: html_document
---

# Write a beta GLM in STAN

```{r}
library(rstan)
model <- "
// beta Regression
data {
  int<lower=1> N;                   // sample size
  int<lower=1> p;                   // p predictors
  vector<lower=0,upper=1>[N] y;     // response 
  matrix[N, p] X;                   // predictor matrix
}
parameters {
  vector[p] theta;                  // reg coefficients
  real<lower=0> phi;                // dispersion parameter
}
transformed parameters{
  vector[p] beta;
  vector[N] A;             // parameter for beta distn
  vector[N] B;             // parameter for beta distn
  vector[N] mu;    // transformed linear predictor
  vector[N] Xbeta;                  // linear predictor
  beta <- theta * 10;               // matt trick if desired
//  vector<lower=0,upper=1>[N] mu;    // transformed linear predictor
  Xbeta <- X * beta;
  for (i in 1:N) { 
    mu[i] <- inv_logit(Xbeta[i]);   
  }
  A <- mu * phi;
  B <- (1.0 - mu) * phi;
}
model {
  // model calculations
  // priors
  theta ~ normal(0, 1);   
  phi ~ cauchy(0, 5);               // different options for phi  
  //phi ~ inv_gamma(.001, .001);
  //phi ~ uniform(0, 500);          // put upper on phi if using this
  // likelihood
  y ~ beta(A, B);
} generated quantities {
vector[N] log_lik;
  for (n in 1:N){
    log_lik[n] <- beta_log(y[n], A[n], B[n]);
  }
}
"
write(model, file = "~/mysis/stanModels/betaRegression-glm.stan")
sm <- stan_model("~/mysis/stanModels/betaRegression-glm.stan")
```

## Simulate some Data

```{r}
library(betareg)
set.seed(123)
N = 80
beta <- c(2, -1, 1, -2, -1)
x0 <- rep(1, N)
x1 <- c(rep(1, N/2), rep(0, N/2)) ## Indicator for net size
x2 <- rep(c(rep(1, 10), rep(0, 30)), 2) ## Indicator for July
x3 <- rep(c(rep(0, 10), rep(1, 20), rep(0, 10)), 2) ## Indicator for August
# x4 <- rep(c(rep(0, 30), rep(1, 10)), 2) ## Indicator for September
x12 <- x1 * x2 ## Interaction between net size and July
x13 <- x1 * x3 ## Interaction between net size and August
# X <- cbind(x0, x1, x2, x3, x12, x13)
X <- cbind(x1, x2, x3, x12, x13)

mu = plogis(X%*%beta)  # add noise if desired + rnorm(N, sd=.01)
phi = 10
A = mu*phi
B = (1-mu)*phi
y = rbeta(N, A, B)
layout(matrix(c(1, 1, 2, 3), 2, 2, byrow=TRUE))
hist(y, "FD")
hist(y[x1 == 0], "FD")
hist(y[x1 == 1], "FD")
```

@@

```{r}
m <- sampling(sm, data = list(N = N, y = y, X = X, p=dim(X)[2]),
  pars = c("beta", "phi"),
  iter = 5000, chains = 4)
```

@@

```{r}
e <- rstan::extract(m, pars = c("beta", "phi"))
true_pars <- c(beta = beta, phi = phi)
library(betareg)
brmod <- betareg(y ~ . -1, data = data.frame(y, X))
summary(brmod)
coefs <- c(brmod$coefficients$mean, brmod$coefficients$precision)
par(mfrow = c(2, 3))
for(i in 1:2) {
  if(i == 1) {
#     for(j in 1:6) {
    for(j in 1:5) {
      plot(density(e[[i]][,j]), main = names(true_pars)[i + j - 1])
      abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
      abline(v = coefs[i + j - 1], lwd = 3, col = "red")
    }
  } else if(i == 2) {
    plot(density(e[[i]]), main = names(true_pars)[i + j - 1])
    abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
    abline(v = coefs[i + j - 1], lwd = 3, col = "red")
  }
}
# legend("topright", legend = c("posterior", "true", "MASS::glm.nb"),
#   col = c("black", "grey", "red"), lty = c(1, 2, 1), lwd = c(1, 1.3, 3),
#   bty = "n")
```

@@

idx <- order(colMeans(e$beta))

ggs_caterpillar(S, family = "beta") + 
  geom_point(aes(y=1:5, x=beta[idx]), col = "red")
```

```{r}
hist(y[X[, 1] == 0])
hist(y[X[, 1] == 1])
```

>>>>

---
title: "Beta Regression"
author: "John Tipton"
date: "August 14, 2015"
output: html_document
---
# Write a beta GLM in STAN

```{r}
library(rstan)
sm <- stan_model("~/mysis/stanModels/betaRegression-glm.stan")
```

## Load the Data

```{r}
library(ggplot2)
set.seed(123)
data_one_way <- read.csv("~/mysis/data/ttestData.csv", skip = 3, nrows = 40, header = TRUE, stringsAsFactors=FALSE)

prop <- c(as.numeric(data_one_way$full.2),    ## proportion of females full net
             as.numeric(data_one_way$half.2)) ## proportion of females half net
N <- length(prop)
## Set improper values to NA

date <- rep(as.factor(data_one_way$month), 2)
net <- as.factor(c(rep("full", 40), rep("half", 40)))
data <- data.frame(prop=prop, date=date, net=net)
X <- model.matrix(~ net * date)
ggplot(data=data, aes(prop)) + geom_histogram() + facet_grid(net ~ .)
```

@@

```{r}
m <- sampling(sm, data = list(N=N, y = prop, X = X, p=dim(X)[2]),
  pars = c("beta", "phi"),
  iter = 5000, chains = 4)
```

@@

```{r}
e <- rstan::extract(m, pars = c("beta", "phi"))
library(betareg)
brmod <- betareg(prop ~ . -1, data = data.frame(prop, X))
summary(brmod)
coefs <- c(brmod$coefficients$mean, brmod$coefficients$precision)
par(mfrow = c(3, 3))
for(i in 1:2) {
  if(i == 1) {
    for(j in 1:6) {
    # for(j in 1:5) {
      plot(density(e[[i]][,j]), main = paste("beta", j, sep=""))
      # abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
      abline(v = coefs[i + j - 1], lwd = 3, col = "red")
    }
  } else if(i == 2) {
    plot(density(e[[i]]), main = "phi")
    # abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
    abline(v = coefs[i + j - 1], lwd = 3, col = "red")
  }
}
# legend("topright", legend = c("posterior", "true", "MASS::glm.nb"),
#   col = c("black", "grey", "red"), lty = c(1, 2, 1), lwd = c(1, 1.3, 3),
#   bty = "n")
```

@@

ggs_caterpillar(S, family = "beta")

```

### 2015-10-12 lengthModel/lengthModel.Rmd

  // data model:
  y ~ normal(X * beta, s);
}
"
write(model, file = "~/mysis/stanModels/length.stan")
# sm <- stan_model("~/mysis/stanModels/length.stan")
sm <- stan("~/mysis/stanModels/length.stan", 
           data=list(N=length(y), p=dim(X)[2], y=y, X=X),
           chains=0)
```
## Fit the Model

>>>>

  // data model:
  for(i in 1:N){
    y[i] ~ normal(X[i] * beta, s[idx[i]]);
  }
}
"
write(model, file = "~/mysis/stanModels/length.stan")
# sm <- stan_model("~/mysis/stanModels/length.stan")
sm <- stan("~/mysis/stanModels/length.stan", 
           data=list(N=length(y), p=dim(X)[2], y=y, X=X),
           chains=0)
```
## Fit the Model

### 2015-10-12 results/mysisResults.Rmd

prop <- c(as.numeric(data_one_way$full.2),    ## proportion of females full net
             as.numeric(data_one_way$half.2)) ## proportion of females half net
N <- length(prop)
## Set improper values to NA

>>>>

prop <- c(as.numeric(data_one_way$full.2),    ## proportion of juveniles full net
             as.numeric(data_one_way$half.2)) ## proportion of juveniles half net
N <- length(prop)
## Set improper values to NA
